When, why, and how to avoid, prefer, or tolerate inbreeding

### 2018-04-02 presentation/inbreeding.Rmd
New slides

## Individual-based models: general conclusions

@@

@@

@@

@@

## Individual-based models: general conclusions

<hr>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016.pdf). <br>

</span>

## Individual-based models: general conclusions

<hr>
<span style="font-size: 18px;">
2. Duthie, AB et al. (2016) *Evolution* [**70**:1927-1943](https://bradduthie.github.io/DuthieEtAl2016b.pdf). <br>
</span>

## Individual-based models: general conclusions

<hr>
<span style="font-size: 18px;">
3. Duthie, AB et al. (2018) *J. Evol. Biol.* [**31**:31-45](https://bradduthie.github.io/DuthieEtAl2017.pdf). <br>
</span>

@@

>>>>

## Individual-based models: general conclusions {.smaller}

@@

- Inbreeding depression strong (avoidance) or very weak (preference)
- Low direct costs of non-random mating
- Small population size where risk of inbreeding is high

<hr>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016.pdf). <br>
</span>

## Individual-based models: general conclusions {.smaller}

- Inbreeding depression strong (avoidance) or very weak (preference)
- Low direct costs of non-random mating
- Small population size where risk of inbreeding is high

@@

- Strong constraints on initial mate choice
- Opportunity for post-copulatory inbreeding adjustment

<hr>
<span style="font-size: 18px;">
1. Duthie, AB & JM Reid (2016) *Am. Nat.* [**188**:651-667](https://bradduthie.github.io/Duthie_Reid2016.pdf). <br>
2. Duthie, AB et al. (2016) *Evolution* [**70**:1927-1943](https://bradduthie.github.io/DuthieEtAl2016b.pdf). <br>
</span>

## Individual-based models: general conclusions {.smaller}

- Inbreeding depression strong (avoidance) or very weak (preference)
- Low direct costs of non-random mating
- Small population size where risk of inbreeding is high

- Strong constraints on initial mate choice
- Opportunity for post-copulatory inbreeding adjustment

@@

- Selection for less costly way to avoid inbreeding
- Polyandry modulates importance of pre-copulatory versus post-copulatory inbreeding avoidance

@@

@@

## What parental investment is, specifically

**Anything that a parent does to increase its offspring's viability *at the expense of its other actual or potential offspring*** [1]

<br><br><br>

(Includes, but is not limited to, behaviour)

<br><br><br><br><br><br>

<hr>
<span style="font-size: 18px;">
1. Trivers R. (1974) *Am. Zool.* [**14**:249–264](https://doi.org/10.1093/icb/14.1.249) <br>
</span>

## Why parental investment is important

- Models of inbreeding assume no concurrent modulation of parental investment or offspring production

- Parents might be able to increase the viability of their inbred offspring through parental investment [1]

- Parents are more closely related to inbred offspring than outbred offspring

![](images/N_vespilloides_1.jpg)

</div>
<br>
<hr>
<span style="font-size: 18px;">
1. Pilakouta, N et al. (2015) *PNAS* [**112**:201500658](http://www.pnas.org/content/112/26/8031). <br>
</span>

## Parental investment and offspring number

<div class = "columns-2">
```{r, echo = FALSE, fig.width = 5}
M <- 10;
m <- seq(from = 0, to = 4, by = 0.1);
n <- M/m;
par(mar = c(5, 5, 1, 1));
plot(x = m, y = n, type = "l", lwd = 3, cex.lab = 1.5, cex.axis = 1.5,
     xlab = "Parental investment (m)", ylab = "Offspring produced (n)",
     ylim = c(0, 20));
text(x = 3.25, y = 16, labels = expression(paste(n == frac(M,m))), cex = 2.5);
```

- Offspring produced (n) decreases with investment per offspring (m)

## Parental investment and offspring number

<div class = "columns-2">
```{r, echo = FALSE, fig.width = 5}
V <- 1 * (1 - exp(-1 * (m - 1)));
V[V<0] <- NA;
par(mar = c(5, 5, 1, 1));
plot(x = m, y = V, type = "l", lwd = 3, cex.lab = 1.5, cex.axis = 1.5,
     xlab = "Parental investment (m)", 
     ylab = expression(paste("Offspring viability (",upsilon,")")),
     ylim = c(0, 1));
text(x = 1.2, y = 0.9, labels = expression(upsilon == 1 - e^{-c(m - m[min])}), 
     cex = 1.75);
```

- Offspring produced ($n$) decreases with investment per offspring ($m$)
- Offspring viability ($\upsilon$) increases with investment per offspring

## Parental investment and offspring number

<div class = "columns-2">
```{r, echo = FALSE, fig.width = 5}
V <- 1 * (1 - exp(-1 * (m - 1)));
V[V<0] <- NA;
par(mar = c(5, 5, 1, 1));
plot(x = m, y = V, type = "l", lwd = 3, cex.lab = 1.5, cex.axis = 1.5,
     xlab = "Parental investment (m)", 
     ylab = expression(paste("Offspring viability (",upsilon,")")),
     ylim = c(0, 1));
text(x = 1.2, y = 0.9, labels = expression(upsilon == 1 - e^{-c(m - m[min])}), 
     cex = 1.75);
abline(a = 0, b = 2*0.3410784 / 2.146197, lty = "dashed");
arrows(x0=2.146197,y0=0.3,x1=2.146197,y1=0,
       length = 0.15,angle=30,code=2,lwd=3);
text(x=2.146197,y=0.4,labels=expression(paste(italic(m)^"*")),cex=2);
```

- Offspring produced ($n$) decreases with investment per offspring ($m$)
- Offspring viability ($\upsilon$) increases with investment per offspring
- Highest rate of viable offspring production defines optimal parental investment ($m^{*}$)
